Psychological Needs During Intergroup Contact

### 2021-11-24 manuscript.tex
paragraph comments added

% Relevance
Conflict between social groups and their individual members remains a prevalent feature of the modern human condition. Prejudices, discrimination, and animosities continue to plague societies around the world. One of the main ameliorations proposed by social psychologists has been positive intergroup contact. Based on Allport's (\citeyear{Allport1954b}) contact hypothesis, a plethora of studies and interventions have worked on bringing different social groups together \citep[e.g.,][]{AlRamiah2012a, Reimer2021}. 
% Yet, despite ...

In its most essential interpretation, the intergroup contact hypothesis postulates that frequent and positive contact with an out-group reduces prejudice and increases favorable attitudes towards the other group \citep[e.g.,][]{Hewstone1996, Pettigrew1998}. A key condition for these contact benefits, has been that the interaction is indeed perceived as positive --- making the interaction quality a crucial mechanism of inter-group contact \citep[e.g.,][]{MacInnis2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for a positive contact \citep[Allport's Optimal Contact conditions,][]{Allport1954b, Pettigrew1969}. And indeed a major meta-analytic review showed, that intergroup contact benefits were larger when Allport's conditions were met. However, the meta-analysis also showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why exactly an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects.

In this article, we propose that the fulfillment of fundamental psychological needs might offer a psycho-social mechanism for understanding and explaining contact quality and in extension positive outgroup attitudes. That is to say, that if someone seeks/wishes to ... . 
Additionally, such needs might even explain the effectiveness of Allport's optimal contact conditions, if these conditions constitute common psychological needs of the interacting individual. To test this we collected three sets of real-life data from recent immigrants, tracking situational needs, interaction quality, and outgroup attitudes, following their daily interactions with host majority group members.

@@

As one of the most researched and practically implemented theories of social psychology, intergroup contact has received a vast amount of empirical, and theoretical attention \citep[][]{Dovidio2017}. Several meta-analytic reviews have found moderately sized positive effects of intergroup contact \citep[][]{Tropp2005, Pettigrew2006, Davies2011} and later meta-analyses have additionally shown that intergroup contact is also effective outside the lab \citep[][]{Beelmann2014, Lemmer2015}, during online contact \citep[][]{White2020}, as well as during several more indirect forms of contact \citep[][]{Miles2014, Zhou2019, Harwood2021}. With such a persistent and meaningful effect, researchers have long proposed theoretical models to understand these intergroup contact effects.

Interestingly, one concept that has arguably remained at the center of theoretical debates throughout the decades has been the role of 'positive' contact. In his now famous chapter on 'the effects of contact', Gordon Allport (\citeyear{Allport1954b}) set out to summarize under what circumstances an interaction between groups tends to be positive, leading to favourable attitudes and reduced prejudice. In the years that followed, several additional conditions were proposed under which a positive intergroup contact would occur \citep[for a critical review see][]{Pettigrew1986}. As a counter-movement to this rather restrictive approach, with the beginning of the 21\textsuperscript{st} century attention had shifted towards investigating the psychological mechanisms of intergroup contact \citep[e.g. see,][]{Paolini2021}. With this shift in focus, the idea of 'positive contact' has seemingly seen a renaissance, where the adjective 'positive' is now not merely an evaluation of the outcome, but rather a feature of the interaction itself. As such, recent work has focused on the role of 'perceived interaction quality' \citep[e.g.,][]{Brown2007, Voci2003} or 'interaction valence' \citep[e.g.,][]{Tropp2016, Barlow2012}, to show that only if an interaction is perceived as positive will it lead to more favourable intergroup relations. And inversely interactions that are perceived as negative have the strong propensity to harm outgroup attitudes. However, relatively little research has investigated why interactions are perceived as positive.

Thus although a sizeable number of moderators and mediators of intergroup contact have been proposed, their effects might be undermined if they do not result in a 'positive' contact. As an example, important moderators and mediators that have been identified are different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, there is a discernible disconnect between these mechanisms and perceptions that an interaction is positive. That is to say that, just because someone learns something new about the other group, or reduces their threat perceptions does not mean that such an interaction is perceived as positive. Or inversely, an interaction can, for example, be very intimate but perceived as negative. If that were the case, there is a reasonable chance that such an interaction might not translate into improved outgroup attitudes \citep[e.g.,][]{Barlow2012}. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

@@

\section{Psychological Needs in Intergroup Relations}
Although considerations of psychological needs and motivations have remained markedly absent within the intergroup contact literature, the concepts are in no way foreign to the general topic of intergroup relations. On the most general level, research on intergroup conflicts has, for example, shown that addressing the group-specific needs predicts willingness to reconcile with the other group \citep[][]{Shnabel2008}. For the case of refugee migrants in their relationship with a majority group, a recent study found that identity needs will buffer against discrimination experiences and will protect personal health \citep[][]{Celebi2017}. Additionally, within the field of interpersonal relations, we find first evidence that the fulfillment of fundamental psychological needs might predict perceived interaction quality in non-intergroup interactions \citep[][]{Downie2008}.

And even within the field of intergroup contact the idea of need fulfillment as a psychological mechanism is not entirely novel. \citet{Dovidio2017} in their narrative review have suggested that understanding psychological needs might be essential to constructive intergroup contact. And recently first empirical research has even found psychological need fulfillment during intergroup contacts to predict support for social change  \citep[][]{Hassler2021}. There is thus initial evidence to suggest that need fulfillment during intergroup contact might offer a promising psychological mechanism to understand perceived interaction quality and positive outgroup attitudes.

One reason why motivational considerations might have remained absent from the intergroup contact literature is that people might enter into a contact with a plethora of individual needs or goals that might be relevant to them in that particular situation. Researchers considering the motivational content would, thus, either test few hyper-specific needs that might not be transferable to other intergroup contexts or they may need to assess a broad range of fundamental psychological needs. To avoid this predicament we propose to start by using adaptive and responsive survey designs that allow a tailored approach based on the participants inputs \citep[e.g.,][]{Tourangeau2017}. In particular, we propose to ask the participants to report their main goal during the interaction in a short open-ended question (i.e., situation core need) and with reference to their own response the participants can then indicate how much this need was fulfilled during the interaction (i.e., need fulfillment). Such an adaptive approach allows us to test the fundamental psychological mechanism of whether situational psychological needs indeed predict perceived interaction quality and positive outgroup attitudes.

\section{Real-Life Process Data}
Over the past two decades several reviewers of the field have advocated that to truly understand intergroup contacts, there is a need for studies that collect longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life data outside the lab \citep[][]{MacInnis2015, McKeown2017}. The reasoning behind these calls for longitudinal real-life data has been manifold. Among the most pressing issues is concern that past investigations of intergroup contact have either focused on the dynamics of individual interactions but within artificial laboratory settings or have collected data situated within the lived experiences but used retrospective self-report data that asked about past interactions in general. The collection of data over time would, thus, allow researchers to follow the development of intergroup relations as they develop across multiple interactions. At the same time, such longitudinal data would also be situated within the real-life experiences of the participant. Additionally, such experience-sampling data can be close to the intergroup interactions would largely mitigate recall biases. Moreover, intensive longitudinal data often allows to capture large amounts of high-quality data with relatively few participants.

In the past such data collections were often unfeasible because they were either physically impractical or too expensive. However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}. At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[see, e.g.,][]{ODonnell2021}. Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
\end{enumerate}
Beyond these main hypotheses we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}). Given the more complex data structure of the extensive longitudinal data, we opted to test our hypotheses using a multilevel regression model, where the measurement occasions (level 1) were nested within the participants (level 2). Given that more complex dynamic modeling procedures are not necessary to test these more basic hypotheses, we will focus on the contemporaneous effect of the data (e.g., need fulfillment during the intergroup interaction predicts outgroup attitudes following the interaction). Such an approach is also tolerant to missing data and uneven case numbers within participants.

To test our main hypotheses, we collected three independent sets of extensive longitudinal data (Studies 1–3). The full surveys are available in Supplementary Materials A as well as in our open science repository (including a complete codebook, see OSF-REFERENCE). Our most comprehensive study (Study 3), was fully preregistered and is also available via our open science repository \citep[][]{Kreienkamp2021f}. The full annotated analysis is available in Supplementary Material B. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (e.g., random slopes model; for the full modeling process see Supplementary Material B).

@@

\section{Discussion}
In this paper we set out to (1) test the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data, (2) we proposed that the fulfillment situational psychological needs meaningfully predicts positive outgroup attitudes, and (3) that the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

When considering the results of the three studies jointly we find that for the basic intergroup contact hypothesis we find mixed results. The aggregated number of interactions with the outgroup across all measurements only had a significant effect on average outgroup attitudes in one of the three studies and in that study there was no significant effect of average interaction quality on outgroup attitudes (which was significant for the other two studies). This meant that only for our smallest study did we find the expected interaction between interaction frequency and interaction quality when looking at the aggregated data across all measurements. However, once we considered the effect of having an interaction with an outgroup member (vs. not having an outgroup interaction) within the participants, we find that intergroup contact is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account). This finding is surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). Even though this inconsistency with past research might be a data artefact (e.g., because most people reported more substantially mode measurements during which they did not have an outgroup interaction), the absence of this aggregate effect could underline the fact that cross-sectional retrospective data might suffer from a recall bias (e.g., where times with no outgroup interactions are undervalued by participants during retrospective evaluations). However, coming to the rescue of the basic intergroup contact tenets are the results from the within-person data. Across all studies, we find that after intergroup contacts participants have substantially more positive outgroup attitudes then at times when they did not have an outgroup interaction. In summary, we thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data we find inconsistent results from the past literature but within participants intergroup contacts are significantly related with positive outgroup attitudes.

Using the data from our third study, we find that Allport's conditions are related to higher interaction quality perceptions and more positive outgroup attitudes. And when we consider interaction quality and Allport's conditions jointly, we find further evidence that part of the benefits of Allport's conditions are due to the case that such interactions are also perceived as more positive. We thus find first evidence that Allport's conditions of optimal contact are also relevant to the daily interactions recent migrants have in their interactions with majority group members.

Finally, when looking at the results for our main proposal about the importance of situational key needs we find highly consistent positive results across all three studies. We find that in all three extensive longitudinal datasets, the fulfillment of core needs during intergroup contacts predicts higher interaction quality perceptions and more positive outgroup attitudes. We find that in all three studies the effect of core needs on outgroup attitudes is either fully or partially through quality perceptions, and we find that this is the case even when taking other fundamental psychological needs or Allport's conditions into account. In fact we find that our core situational needs measure predicts outgroup attitudes slightly better than Allport's conditions and consistently explained more variance in outgroup attitudes than any of the other psychological needs. In most cases the core need even takes over the variances previously explained by the self-determination theory needs. We thus find strong evidence that within everyday life interactions of recent migrants with majority group members, the perception that one's interaction specific psychological needs are fulfilled offers a meaningful and flexible predictor of outgroup attitudes. As such the data offers a first test of the motivational basis of intergroup contact.

\subsection{Limitations}
\begin{enumerate}
    \item minority perspective only
    \item lagged effects left for future studies (potentially not relevant, time scale)
    \item ESM: real-life data comes at the expense of having short/single item operationalizations
\end{enumerate}

>>>>

% Relevance
Conflict between social groups and their individual members remains a prevalent feature of the modern human condition. Prejudices, discrimination, and animosities continue to plague societies around the world. One of the main ameliorations proposed by social psychologists has been positive intergroup contact. Based on Allport's (\citeyear{Allport1954b}) contact hypothesis, a plethora of studies and interventions have worked on bringing different social groups together \citep[e.g.,][]{AlRamiah2012a, Reimer2021}.

% Problem
In its most essential interpretation, the intergroup contact hypothesis postulates that frequent and positive contact with an out-group reduces prejudice and increases favorable attitudes towards the other group \citep[e.g.,][]{Hewstone1996, Pettigrew1998}. A key condition for these contact benefits, has been that the interaction is indeed perceived as positive --- making the interaction quality a crucial mechanism of inter-group contact \citep[e.g.,][]{MacInnis2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for a positive contact \citep[Allport's Optimal Contact conditions,][]{Allport1954b, Pettigrew1969}. And indeed a major meta-analytic review showed, that intergroup contact benefits were larger when Allport's conditions were met. However, the meta-analysis also showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why exactly an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects.

% Aim / Solution
In this article, we propose that the fulfillment of fundamental psychological needs might offer a psycho-social mechanism for understanding and explaining contact quality and positive outgroup attitudes. That is to say, that if a person, for example, seeks to feel accepted by their interaction partner and this need is fulfilled during the interaction, the person should rate the interaction and the group of the interaction partner more favorably. Additionally, such needs might even explain the effectiveness of Allport's optimal contact conditions, if these conditions constitute common psychological needs of the interacting individual. To test this we collected three sets of real-life data from recent immigrants, tracking situational needs, interaction quality, and outgroup attitudes, following their daily interactions with majority group members.

@@

% Intergroup contact = Robust and strong effect (from meta-analyses) → much theoretical interest
As one of the most researched and practically implemented theories of social psychology, intergroup contact has received a vast amount of empirical, and theoretical attention \citep[][]{Dovidio2017}. Several meta-analytic reviews have found moderately sized positive effects of intergroup contact \citep[][]{Tropp2005, Pettigrew2006, Davies2011} and later meta-analyses have additionally shown that intergroup contact is also effective outside the lab \citep[][]{Beelmann2014, Lemmer2015}, during online contact \citep[][]{White2020}, as well as during several more indirect forms of contact \citep[][]{Miles2014, Zhou2019, Harwood2021}. With such a persistent and meaningful effect, researchers have long proposed theoretical models to understand these intergroup contact effects.

% what makes a 'positive contact' has received much theoretical interest. (1) conditions of positive contact, then (2) contact quality. But why is it perceived as positive? 
Interestingly, one concept that has arguably remained at the center of theoretical debates throughout the decades has been the role of 'positive' contact. In his now famous chapter on 'the effects of contact', Gordon Allport (\citeyear{Allport1954b}) set out to summarize under what circumstances an interaction between groups tends to be positive, leading to favourable attitudes and reduced prejudice. In the years that followed, several additional conditions were proposed under which a positive intergroup contact would occur \citep[for a critical review see][]{Pettigrew1986}. As a counter-movement to this rather restrictive approach, with the beginning of the 21\textsuperscript{st} century attention had shifted towards investigating the psychological mechanisms of intergroup contact \citep[e.g. see,][]{Paolini2021}. With this shift in focus, the idea of 'positive contact' has seemingly seen a renaissance, where the adjective 'positive' is now not merely an evaluation of the outcome, but rather a feature of the interaction itself. As such, recent work has focused on the role of 'perceived interaction quality' \citep[e.g.,][]{Brown2007, Voci2003} or 'interaction valence' \citep[e.g.,][]{Tropp2016, Barlow2012}, to show that only if an interaction is perceived as positive will it lead to more favourable intergroup relations. And inversely interactions that are perceived as negative have the strong propensity to harm outgroup attitudes. However, relatively little research has investigated why interactions are perceived as positive.

% many moderators and mediators but none of these necessarily leads to higher interaction quality. Solution: look at need fulfillment (motivational literature)
Thus although a sizeable number of moderators and mediators of intergroup contact have been proposed, their effects might be undermined if they do not result in a 'positive' contact. As an example, important moderators and mediators that have been identified are different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, there is a discernible disconnect between these mechanisms and perceptions that an interaction is positive. That is to say that, just because someone learns something new about the other group, or reduces their threat perceptions does not mean that such an interaction is perceived as positive. Or inversely, an interaction can, for example, be very intimate but perceived as negative. If that were the case, there is a reasonable chance that such an interaction might not translate into improved outgroup attitudes \citep[e.g.,][]{Barlow2012}. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

@@

\section{Psychological Needs in Intergroup Relations}
% Important in broader field: needs have been found to be important in broader intergroup relations and interpersonal contact literature.
Although considerations of psychological needs and motivations have remained markedly absent within the intergroup contact literature, the concepts are in no way foreign to the general topic of intergroup relations. On the most general level, research on intergroup conflicts has, for example, shown that addressing the group-specific needs predicts willingness to reconcile with the other group \citep[][]{Shnabel2008}. For the case of refugee migrants in their relationship with a majority group, a recent study found that identity needs will buffer against discrimination experiences and will protect personal health \citep[][]{Celebi2017}. Additionally, within the field of interpersonal relations, we find first evidence that the fulfillment of fundamental psychological needs might predict perceived interaction quality in non-intergroup interactions \citep[][]{Downie2008}.

% Promising for intergroup contact: Asked for in intergroup contact reviews and first research looked at social change
And even within the field of intergroup contact the idea of need fulfillment as a psychological mechanism is not entirely novel. \citet{Dovidio2017} in their narrative review have suggested that understanding psychological needs might be essential to constructive intergroup contact. And recently first empirical research has even found psychological need fulfillment during intergroup contacts to predict support for social change \citep[][]{Hassler2021}. There is thus initial evidence to suggest that need fulfillment during intergroup contact might offer a promising psychological mechanism to understand perceived interaction quality and positive outgroup attitudes.

% Difficult to test general mechanism: Either consider too many needs (situationally relevant but not feasible) or too few (feasible but not transferable). Solution: ask for main need + rating of main need  
One reason why motivational considerations might have remained absent from the intergroup contact literature is that there is an overwhelming number of individual needs or goals that might be relevant to a person during an intergroup interaction. Researchers considering the motivational content would, thus, either test few hyper-specific needs that might not be transferable to other intergroup contexts or they may need to assess a broad range of fundamental psychological needs. To avoid this predicament we propose to start by using adaptive and responsive survey designs that allow a tailored approach based on the participants inputs \citep[e.g.,][]{Tourangeau2017}. In particular, we propose to ask the participants to report their main goal during the interaction in a short open-ended question (i.e., situation core need) and with reference to their own response the participants can then indicate how much this need was fulfilled during the interaction (i.e., need fulfillment). Such an adaptive approach allows us to test the fundamental psychological mechanism of whether situational psychological needs indeed predict perceived interaction quality and positive outgroup attitudes.

\section{Real-Life Process Data}
% call for longitudinal and real-life data b/c: lab process test vs. real-life retrospective. + less recall bias + much data from few PPTs
Over the past two decades several reviewers of the field have advocated that to truly understand intergroup contacts, there is a need for studies that collect longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life data outside the lab \citep[][]{MacInnis2015, McKeown2017}. The reasoning behind these calls for longitudinal real-life data has been manifold. Among the most pressing issues is concern that past investigations of intergroup contact have either focused on the dynamics of individual interactions but within artificial laboratory settings or have collected data situated within the lived experiences but used retrospective self-report data that asked about past interactions in general. The collection of data over time would, thus, allow researchers to follow the development of intergroup relations as they develop across multiple interactions. At the same time, such longitudinal data would also be situated within the real-life experiences of the participant. Additionally, such experience-sampling data can be close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, intensive longitudinal data often allows to capture large amounts of high-quality data with relatively few participants.

% Used to be difficult but technological and methodological developments (e.g., FormR + HLM) → collect large body of intensive longitudinal data
In the past such data collections were often unfeasible because they were either physically impractical or too expensive. However, recent technological developments allow us to easily collect experience sampling data on mobile devices \citep[e.g.,][]{Keil2020} or using web-based applications \citep[e.g.,][]{Arslan2020}. At the same time, analytical methods for such more complex data have become more readily available, making the analyses more approachable \citep[see, e.g.,][]{ODonnell2021}. Given these, technological and methodological developments, we were able to collect a large amount of real-life data following the daily intergroup interactions of recent migrants with the majority group members.

\section{The Present Research}
The aim of this paper is essentially threefold. First, our study is among the first to test the fundamental tenets of intergroup contact and Allport's conditions in real-life extensive longitudinal data. Second, we test whether the fulfillment situational psychological needs is meaningfully related to more positive outgroup attitudes following intergroup interactions. And third, we test the role of interaction quality in  the relationship of psychological need fulfillment and outgroup attitudes. Based on this we formulated three main hypotheses:
\begin{enumerate}[label=H\arabic*:]
    \item Based on the most general understanding of the contact hypothesis, an increase in frequency and quality of contact should jointly account for changes in more favorable outgroup attitudes across extensive longitudinal data.
    \item Based on Allport’s optimal contact conditions, intergroup interactions with equal status, common goals, collaboration, and structural support should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
    \item Based on our proposal, intergroup interactions with higher situational core need fulfillment should predict more favorable outgroup attitudes due to more positive interaction quality perceptions within the extensive longitudinal data.
\end{enumerate}
Beyond these main hypotheses we have formulated several sub-hypotheses that form the basis for our analysis plan (these sub-hypotheses also include several robustness checks and are available in Appendix \ref{app:AppendixHypotheses}). Given the more complex data structure of the extensive longitudinal data, we opted to test our hypotheses using a multilevel regression model, where the measurement occasions (level 1) were nested within the participants (level 2). Given that more complex dynamic modeling procedures are not necessary to test these more basic hypotheses, we will focus on the contemporaneous effect of the data (e.g., need fulfillment during the intergroup interaction predicts outgroup attitudes following the interaction). Such an approach is also tolerant to missing data and uneven case numbers within participants.

To test our main hypotheses, we collected three independent sets of extensive longitudinal data (Studies 1–3). The full surveys are available in Supplementary Materials A as well as in our open science repository (including a complete codebook, see OSF-REFERENCE). Our most comprehensive study (Study 3), was fully preregistered, which is also available via our open science repository \citep[][]{Kreienkamp2021f}. The full annotated analyses are available in Supplementary Material B. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (e.g., random slopes model; for the full modeling process see Supplementary Material B).

@@

\section{Discussion}
% aims re-iterated
In this paper we set out to (1) test the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data, (2) we proposed that the fulfillment situational psychological needs meaningfully predicts positive outgroup attitudes, and (3) that the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we find mixed results for the basic intergroup contact hypothesis. The aggregated number of outgroup interactions across all measurements only had a significant effect on average outgroup attitudes in one of the three studies and in that study there was no significant effect of average interaction quality on outgroup attitudes (which was significant for the other two studies). This meant that only for our smallest study did we find the expected interaction between interaction frequency and interaction quality when looking at the aggregated data across all measurements. However, once we considered the effect of having an interaction with an outgroup member (vs. not having an outgroup interaction) within the participants, we find that intergroup contact is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account). While this later results is in line with our expectations, the earlier aggregate findings are surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). Even though this inconsistency with past research might be a data artefact (e.g., because most people reported more substantially mode measurements during which they did not have an outgroup interaction), the absence of this aggregate effect could underline the fact that cross-sectional retrospective data might suffer from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations). In summary, we thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data we find inconsistent results from the past literature but within participants intergroup contacts are significantly related with positive outgroup attitudes.

% Allport's conditions in ESM data [consistent predictor and partially through quality]
Using the data from our third study, we find that Allport's conditions are related to higher interaction quality perceptions and more positive outgroup attitudes. And when we consider interaction quality and Allport's conditions jointly, we find further evidence that part of the benefits of Allport's conditions are due to the case that such interactions are also perceived as more positive. We thus find first evidence that Allport's conditions of optimal contact are also relevant to the daily interactions recent migrants have in their interactions with majority group members.

% Need fulfillment in intergroup contact [consistent predictor, through quality, slightly better than Allport and SDT]
Finally, when looking at the results for our main proposal about the importance of situational key needs, we find highly consistent positive results across all three studies. We find that in all three extensive longitudinal datasets, the fulfillment of core needs during intergroup contacts predicts higher interaction quality perceptions and more positive outgroup attitudes. We find that in all three studies the effect of core need fulfillment on outgroup attitudes is either fully or partially through its effect on quality perceptions, and we find that need fulfillment is an important predictor even when taking other fundamental psychological needs or Allport's conditions into account. In fact we find that our core situational needs measure predicted outgroup attitudes slightly better than Allport's conditions and consistently explained more variance in outgroup attitudes than any of the other psychological needs. In most cases the core need even took over the variances previously explained by the self-determination theory needs. We thus find strong evidence that within everyday life interactions of recent migrants with majority group members, the perception that one's interaction specific psychological needs are fulfilled offers a meaningful and flexible predictor of outgroup attitudes. As such the data offers a first test of the motivational basis of intergroup contact.

\subsection{Limitations}
to be discussed, potentially:
\begin{enumerate}
    \item minority perspective only
    \item lagged effects left for future studies (potentially not relevant, time scale)
    \item ESM: real-life data comes at the expense of having short/single item operationalizations
\end{enumerate}
